### 2020-11-09 6_Conclusion.tex

\section{Conclusions}
\label{conclusion}

\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}

{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}