### 2021-01-20 dataset/analysis.Rmd
adicionando os datasets

---
title: "atoms-of-confusion"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(".")
require(sqldf)
```

## Load the Dataset

```{r}
ds <- read.csv("survey-results.csv")
colnames(ds)
nrow(ds)
```

## Clean up the Dataset

```{r}
ds <- ds[ds$ref=="reddit" & ds$total == "YES",]
ds.count <- tapply(ds$time,ds$rid,length)
ds.full <- names(ds.count)[ds.count==20]
finalds <- ds[is.element(el = ds$rid, set = ds.full),]
dim(finalds)
```

## Number of Latin squares

```{r}
sqldf("select count(distinct rid) from finalds")
```

## Effect of atoms of confusion in correctness

```{r}
prop.table(table(finalds$correct,finalds$atom),margin = 2)
chisq.test(finalds$correct,finalds$atom)
```