### 2021-01-20 dataset/survey-results-final.Rmd

---
title: "survey-analysis-final"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='images/', dev=c('png', 'pdf'))

library(questionr)
library(fmsb)
library(sqldf)
library(xtable)
library(effsize)
setwd(".")

```

## Load the datasets

```{r cars}
ds <- read.csv("survey-results.csv")
nrow(ds)
colnames(ds)

atomId <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 0) 
atomDescription <- c("Arithmetic as Logic", "Assignment as Value", "Automatic Semicolon Insertion",
           "Comma Operator", "Ternary Operator", "Implict Predicate", 
           "Logic as Control Flow", "Ommitted Curly Braces", "Post Increment", 
           "Pre Increment")

atom_ds <- data.frame(atomId, atomDescription)
```

## Filter the datasets

```{r}
subSet <- ds[ds$ref=="reddit" & ds$total == "YES",]
summary(subSet$total)
squareLengths <- tapply(subSet$time,subSet$rid,length)
completeCases <- names(squareLengths)[squareLengths==20]
ds <- ds[is.element(el = ds$rid, set = completeCases),]
dim(ds)
```

## Exploratory Data Analysis

### Demographics

#### Education

```{r dem-education}
educationLevelIds <- c(1, 2, 3, 4, 5)
educationLevelLabels <- c("High school degree or equivalent", "Some university course but not a degree", "Bacharelor degree", "Master degree", "Doctor degree")

education_ds <- data.frame(educationLevelIds, educationLevelLabels)
colnames(education_ds) <- c("id", "description")

tmp <- sqldf("select distinct sid, education from ds")

demEducation <- sqldf("select b.description, count(*) total 
                      from tmp a, education_ds b
                      where a.education = b.id
                      group by b.description 
                      order by 2 desc")

barplot(demEducation$total, col=topo.colors(5), ylim = c(0, 70), cex=0.7, cex.lab = 0.7, cex.axis=0.7, ylab="Number of participants")

legend("topright", legend=demEducation$description, fill=topo.colors(5), cex=0.6)
```

#### Experience

```{r dem-experience}
experienceLevelIds <- c(1, 2, 3, 4)
experienceLevelLabels <- c("Under one year of experience", "One year and under four years of experience", "Four years and under ten years of experience", "More than ten years of experience")

experience_ds <- data.frame(experienceLevelIds, experienceLevelLabels)

colnames(experience_ds) <- c("id", "description")

tmp <- sqldf("select distinct sid, experience from ds")
demExperience <- sqldf("select experience, count(*) total from tmp group by experience order by 1")

demExperience <- sqldf("select a.experience, b.description, a.total
                        from demExperience a, experience_ds b
                        where a.experience = b.id 
                        order by 3 desc")
demExperience

barplot(demExperience$total, col=topo.colors(4), ylim = c(0, 60), cex=0.7, cex.lab = 0.7, cex.axis=0.7, ylab="Number of participants")

legend("topright", legend=demExperience$description, fill=topo.colors(4), cex=0.6)
```

### Total number of correct answers (Table III)

```{r correct-answers}
codeWithAtoms <- sqldf("select qid, count(*) confuseCode, avg(time) timeConfuseCode
                        from ds 
                        where atom == 'YES' and correct = 'CORRECT'
                        group by qid")

codeWithoutAtoms <- sqldf("select qid, count(*) cleanCode, avg(time) timeCleanCode
                        from ds 
                        where atom == 'NO' and correct = 'CORRECT'
                        group by qid")

codeWithAtoms["atomId"] = codeWithAtoms$qid %% 10
codeWithoutAtoms["atomId"] = codeWithoutAtoms$qid %% 10

merged <- sqldf("select c.atomDescription as Atom, 
                        a.confuseCode as 'Confuse Code', b.cleanCode as 'Clean Code',  
                        (b.cleanCode * 100 / a.confuseCode) -100 as 'Delta (%)'
                 from codeWithAtoms a, codeWithoutAtoms b, atom_ds c
                 where a.atomId = b.atomId and a.atomId = c.atomId
                 order by 4 desc")

xtable(merged)
```

```{r wrong-answers-plot}
wrongAnswers = ds[ds$correct == 'WRONG', ]

wrongAnswersByStudentTreatment <- aggregate(rid~sid+atom, 
                                            data = wrongAnswers, 
                                            FUN=length)

boxplot(wrongAnswersByStudentTreatment$rid ~ wrongAnswersByStudentTreatment$atom
        , ylab = "", xlab = "", main = "", 
        names = c("Clean Code", "Confuse Code"))
```

### Average time for correct answers (Table IV)

```{r time}
merged <- sqldf("select c.atomDescription as Atom, 
                        a.timeConfuseCode as 'Confuse Code', b.timeCleanCode as 'Clean Code',
                        (b.timeCleanCode * 100 / a.timeConfuseCode) -100 as 'Delta (%)'
                 from codeWithAtoms a, codeWithoutAtoms b, atom_ds c
                 where a.atomId = b.atomId and a.atomId = c.atomId
                 order by 4")  
xtable(merged)

correctAnswers = ds[ds$correct == 'CORRECT', ]

correctAnswersByStudentTreatment <- aggregate(time~sid+atom, 
                                            data = correctAnswers, 
                                            FUN=mean)

boxplot(log(correctAnswersByStudentTreatment$time) ~ correctAnswersByStudentTreatment$atom
        , ylab = "", xlab = "", main = "Number of wrong answers of each participant", 
        names = c("Clean Code", "Confuse Code"))

```

```{r}
prop.table(table(ds$correct,ds$atom),margin = 2)
chisq.test(ds$correct,ds$atom)
```

## Regression Analysis (correctness~atom + experience + education)

```{r}
experience <- as.factor(ds$experience)
education <- as.factor(ds$education) 
contrr<-matrix(c(rep(1,4),c(1/2,1/2,-1/2,-1/2),c(1,-1,0,0),c(0,0,1,-1)),byrow=TRUE,nrow=4)
contrasts(experience)<-solve(contrr)[,2:4]

ds$atom <- as.factor(ds$atom)
ds$correct <- as.factor(ds$correct)

mod <- glm(ds$correct~ds$atom+experience+education,family = "binomial")
summary(mod)
```

## Impact considering Developer Experience

```{r experience}
for(i in 1:4){
  print(paste("Experience ", i ))
  print(prop.table(table(ds$correct[ds$experience==i],ds$atom[ds$experience==i]),margin=2))
  print(chisq.test(table(ds$correct[ds$experience==i],ds$atom[ds$experience==i])))
}
```

## Impact considering Developer Education

```{r education}
for(i in 1:5){
  print(paste("Education ", i ))
  print(prop.table(table(ds$correct[ds$education==i],ds$atom[ds$education==i]),margin=2))
  print(chisq.test(table(ds$correct[ds$education==i],ds$atom[ds$education==i])))
}
```

## Impact of Individual Atoms

```{r}

chiTest <- c()
oddsRatio <- c()
mannWhitneyTest <- c()
cliffDelta <- c()

for(i in 1:10){
  subSet <- ds[is.element(el = ds$qid, set = c(i,i+10)),]
  print(atomDescription[i])

  tableCorrectness <- table(subSet$correct,subSet$atom)
  tableTime <- aggregate(ds$time, by=list(atom=ds$atom), FUN=sum)

  print(tableCorrectness)
  print(tableTime)

  test <- chisq.test(tableCorrectness)
  print(test)

  chiTest[i] <- format.pval(test$p.value) 
  oddsRatio[i] <- odds.ratio(tableCorrectness)$OR

  mannWhitneyTest[i] <- format.pval(wilcox.test(subSet$time~as.factor(subSet$atom))$p.value)
  cliffDelta[i] <- cliff.delta(subSet$time~as.factor(subSet$atom))$estimate

  experience <- as.factor(subSet$experience)
  contrr<-matrix(c(rep(1,4),c(1/2,1/2,-1/2,-1/2),c(1,-1,0,0),c(0,0,1,-1)),byrow=TRUE,nrow=4)
  contrasts(experience)<-solve(contrr)[,2:4]

  mod <- glm(subSet$correct ~ subSet$atom +  experience,family="binomial")
  print(summary(mod))
  print(prop.table(table(subSet$correct,subSet$experience),margin = 2))
}

analysis_df <- data.frame(atomDescription, chiTest, oddsRatio, mannWhitneyTest, cliffDelta)
analysis_df <- analysis_df[order(-oddsRatio),]

colnames(analysis_df) <- c("Atom", "ChiTest", "Odds Ratio Correctness", "Wilcox Test (Time)", "Cliff Delta")

analysis_df

xtable(analysis_df)
```