### 2021-04-08 5_Threats.tex
ajustes nas issues apontadas por Marcio

\section{Threats to Validity}
\label{threat}

%% Since our research was conducted online with unknown participants, and thus we had no way of confirming their
%% levels of education and experience. Also,  we did not have a way to prevent respondents from cheating,
%% such as running the code on an interpreter, or consulting other people. This presents a threat to the validity to the
%% conclusions of our work, even though we leverage some means to try to mitigate this threat (as discussed
%% in Section~\ref{method}).

%% Another threat to our results lies in the fact that some constructs
%% might not be frequent, or present at all, in JavaScript code bases.
%% The atom candidates Arithmetic as Logic and
%% the Comma Operator were rarely found in the
%% open source JavaScript code we mined in our
%% third study.

%% A potential problem with our method is that
%% there little incentive for subjects to think
%% thoroughly about the questions. We observed a lack of
%% engagement when we ran the survey with undergraduate students during
%% a pilot study. Although our final subjects were voluntarily
%% partaking in the survey, we could not be sure that, after some
%% time taking the survey, respondents would become tired and stop thinking
%% clearly about the code.

%% Finally, in one of the atoms, namely the Omitted Curly Braces,
%% we intentionally removed indentation from the original code,
%% which is highly unusual, given that many programmers use automatic
%% formatting in their code editors. This can introduce some level
%% of artificiality to this atom's question. Nonetheless,
%% we discovered that JavaScript does allow the programmer
%% to omit the curly braces after \textit{if} statements,
%% and insert multiple statements in the following line.
%% This fact itself might constitute
%% a source of confusion, which we leave to analyse
%% in our future endeavors.