### 2022-03-29 dataset/survey.Rmd

---
title: "survey-analysis-final"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='images/', dev=c('png', 'pdf'))

library(questionr)
library(fmsb)
library(sqldf)
library(xtable)
library(effsize)
setwd(".")

```

## Load the datasets

```{r}
ds <- read.csv("interview-summary.csv", head=T, sep=";")

meanValues = c("", mean(ds$Left), mean(ds$Right), mean(ds$Neutral))

colnames(ds) <- c("Atom", "LHS (without AOC)", "RHS (with AOC", "Neutral")

ds <- rbind(ds, meanValues)
ds

xtable(ds)
```