### 2022-09-23 declaration.tex
adding letters

\documentclass{letter}
\usepackage{hyperref}
\signature{Rodrigo Bonif\'{a}cio}
\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil}
\begin{document}

\begin{letter}{Journal of Systems and Software Editors-in-Chief}

\opening{Dear Profs. P. Avgeriou and D. Shepherd}

On behalf of the authors of the
paper ``\emph{An Investigation of Confusing Code Patterns in JavaScript}'',
I declare that we have no known competing financial interests or personal relationships
that could have appeared to influence the work reported in this paper.

\closing{Yours Faithfully,}

\end{letter}
\end{document}