### 2022-09-23 letter.tex

\documentclass{letter}
\usepackage{hyperref}
\signature{Rodrigo Bonif\'{a}cio}
\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil}
\begin{document}

\begin{letter}{Journal of Systems and Software Editors-in-Chief}

\opening{Dear Profs. P. Avgeriou and D. Shepherd}

I am writing to you on behalf of the authors of the
paper ``\emph{An Investigation of Confusing Code Patterns in JavaScript}''.
The paper presents the results of an original study that explores the impact of atoms of confusion in
JavaScript code. In more details, we report the results of a mixed methods research
that includes two experiments with 210 JavaScript developers, an interview
study with 15 JavaScript developers, and a software repository mining effort.
The highlights of the paper are as follows:

\begin{itemize}
 \item The experiments confirm that two code patterns previously validated as atoms of confusion for C and C++ languages
  also confuse JavaScript programmers: The Comma Operator and Assignment as Value patterns.

 \item The code patterns Pre- and Post-increment, Omitted Curly Braces, Change of Literal
   Encoding, Implicit Predicate, and Type Conversions have been confirmed as confusing in at least one of our experiments.

 \item Code patterns that we validated as confusing frequently appear in open source programs. In particular,
   we found 19.89 occurrences of the atom Implicit Predicate per KLOC.

 \item Our experiments did not confirm that the patterns Logic as Control Flow and Conditional Operator are confusing for
   JavaScript developers, contrasting with previous research works that present evidence these code patterns are atoms of confusion for C and Java.

 \item Practitioners also consider JavaScript specific idioms (such as Property Access via Array Subscription and Arrow Functions)
   confusing---they are possible atom candidates that deserve further research for validation. 
\end{itemize}

We hope our paper is within the scope of the Journal of Systems and Software and
we kindly ask you to consider it for review.

\closing{Yours Faithfully,}

\end{letter}
\end{document}