### 2022-12-14 package/s03_mining_software_repositories/analysis.Rmd

---
title: "Atoms JS - Prevalence Study"
author: "Rodrigo Bonifácio et al."
date: '2022-12-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(".")
require(sqldf)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
prevalence <- read.csv("dataset.csv")
repositories <- read.csv("repositories.csv")

##  [1] "language"         "id"               "full_name"        "name"            
##  [5] "size"             "stargazers_count" "forks"            "created_at"      
##  [9] "git_url"          "num_lines"        "num_contributors"

colnames(repositories)
colnames(prevalence)
nrow(prevalence)

## [1] "atom"      "atom_name" "full_name" "location"  "url"

sqldf("select atom_name, count(*) 
       from prevalence 
       group by atom_name 
      order by 2 desc")

sqldf("select count(distinct full_name) 
       from prevalence")

sqldf("select atom_name, count(distinct full_name) * 100 / 72.00 as Prevalence
      from prevalence 
       group by atom_name 
       order by 2 desc")

summary(repositories$num_lines)

full_ds <- sqldf("select r.name, r.size, r.num_lines, p.* 
                  from repositories r, prevalence p 
                  where r.full_name = p.full_name")

sqldf("select count(distinct name) from full_ds")

sqldf("select name, size, num_lines, count(distinct atom_name), count(*)
       from full_ds 
       group by name, size, num_lines 
       order by 3 desc")

##                      atom_name Prevalence
## 1              TernaryOperator  100.00000
## 2            ImplicitPredicate  100.00000
## 3           OmittedCurlyBraces   91.66667
## 4      PostIncrementExpression   90.27778
## 5        PreIncremenExpression   83.33333
## 6            AssignmentasValue   81.94444
## 7           LogicasControlFlow   56.94444
## 8  AutomaticSemicolonInsertion   33.33333
## 9            ArithmeticasLogic   23.61111
## 10               CommaOperator   15.27778

sqldf("select * from 
       full_ds where  name = 'carbon' and atom_name = 'ImplicitPredicate'
       order by RANDOM() 
       LIMIT 100")
```