### 2022-12-20 biography.tex
adding paper's metadata

\documentclass{article}

\usepackage[brazil]{babel}
\usepackage[utf8]{inputenc}

\title{Authors Biography}

\begin{document}

\maketitle

\begin{itemize}

\item Adriano Torres is a PhD student at The University of Adelaide, Australia. His alma mater is the University of Brasilia, where he completed his Bachelors and MsC. His research focuses on the development of methods and tools that automate the discovery and correction of software faults. He also has industry experience with distributed systems, and currently works as engineering lead for Luma Health Inc.

\item M\'{a}rcio Okimoto is a Database Administrator and junior researcher. He's a specialist in Business Intelligence and Data Warehousing. He also has a Master's Degree in Computer Science from the University of Brasília (UnB) and his major research interest is the Mining Software Repositories (MSR) field with a focus on productivity.

\item Diego Marc\'{i}lio is a Ph.D. student in the Software institute at USI Università della Svizzera italiana in Lugano, Switzerland.
He received his M.S. in Informatics from the University of Brasília, Brazil, in 2019. His research interests center around aiding developers when understanding, writing, evolving, and maintaining source code.

\item Caio Oliveira has a Computer Science Bachelor's Degree from the University of Brasilia. He is a front-end focused developer with nearly 7
  years of experience with Angular. He has worked in projects as full-stack developer with technologies such as JavaEE, Django, Flask, Spring boot, Laravel, and all of
  them had Angular as the frontend technology. Has familiarity with Javascript and Typescript languages.

\item Fernando Castor is an assistant professor at the Department of Information and Computing Sciences, Utrecht University, The Netherlands,  working under prof. Sjaak Brinkkemper, and an associate professor (on leave) at the Informatics Center, Federal University of Pernambuco, Brazil. His research interests center around program understanding, energy consumption, and exception handling.

\item Rodrigo Bonif\'{a}cio is an associate professor at the Computer Science Department, University of Brasília, Brazil, and a member of the Software Productivity Group, led by Prof. Paulo Borba. His research interests lie in understanding
  the common usage of programming language features, program transformations, software architecture and modularity, and software security.

\item M\'{a}rcio Ribeiro is an associate professor at the Computing Institute, Federal University of Alagoas, Brazil. His research interests lie in program refactoring, program understanding, program testing, software modularity, and high configurable systems.

\item Edna Dias Canedo is currently an associate professor at University of Brasilia (UnB - Brazil). She received her Ph.D. (2012) degree in Electrical Engineering from the University of Brasilia (UnB).
  Her current research interests include Software Engineering, Requirements Engineering, Gender Diversity, Software Systems, Empirical Software Engineering, and Usability.

\item Eduardo Monteiro is an associate professor at the Statistics Department in Universidade de Brasília, Brazil, and focuses his research in statistical modeling and applications.  
\end{itemize}

\end{document}