### 2022-12-20 contributions.tex

\documentclass{article}

\usepackage[brazil]{babel}
\usepackage[utf8]{inputenc}

\title{Authors Contributions}

\begin{document}

\maketitle

\begin{itemize}

\item {\bf Adriano Torres} and {\bf Caio Ribeiro} led all steps in the second study (experiment using Latin Square design). {\bf Adriano Torres} also contributed to the third study (interviews),
  under the supervision of {\bf Edna Dias Canedo}.

\item {\bf M\'{a}rcio Okimoto} and {\bf Diego Marc\'{i}lio} led all steps in the fourth study (MSR effort), implementing the CodeQL queries---under the
  supervision of {\bf Rodrigo Bonif\'{a}cio} and {\bf Edna Dias Canedo}.

\item {\bf Fernando Castor} and {\bf Pedro Queiroga} led all steps in the first study (experiment using the Repeated Measures design). {\bf Fernando Castor}, {\bf Rodrigo Bonif\'{a}cio}, {\bf M\'{a}rcio Ribeiro}, {\bf Edna Dias Canedo}, and {\bf Eduardo Monteiro} co-supervised the whole research.

\item {\bf Eduardo Monteiro} contributed to defining the statistical methods used in this research and implementing several R scripts for evaluating the outcomes of this research.

\item {\bf All authors} have contributed to the writting of the manuscript.    
\end{itemize}

\end{document}