### 2022-12-20 highlights.tex

\documentclass{article}

\title{Paper Highlights}

\begin{document}

\maketitle

\begin{itemize}

\item We present the results of four empirical studies that explores the impact and
  prevalence of atoms of confusion in JavaScript.

\item We bring evidence that two code patterns that have been previously observed to confuse C
  programmers also confuse JavaScript programmers: the comma operator and assignments being used as values.

\item We reveal that some JavaScript-specific elements, such as automatic semicolon insertion and object destructuring,
  also have the potential to cause confusion.

\item We shhow that examined code patterns are frequent and used intensively in 72 popular open-source JavaScript projects:
  four atom candidates appear in 90\% of them and two of them occur more than once for every 100 lines of code.

\end{itemize}

\end{document}