### 2023-01-11 introduction.tex

\section{Introduction}
\label{intro}

% Source code misunderstandings
% Recent work... For instance, consider the pair...
% However, focus on C, C++ and Java.
% In this paper ==> Javascript!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INICIO INTRODUCAO - MARCIO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Understanding code is an expensive and important task. In real-world software development, it is commonplace for developers to spend more time understanding code than actually writing it~\cite{CitationRequired}.Figuring out what parts of an existing system do by reading its source code is necessary for important software development tasks such as testing, debugging, and adding new functionality to a system.

Developers are often confused while reading unfamiliar code~\cite{Ebert:2021:ESC}. According to Hermans~\cite{ProgrammersBrain}, that confusion can stem from three sources: (i) lack of knowledge, i.e., not knowing what an element in the program does; (ii) lack of information, i.e., not knowing how a program element works; and (iii) lack of processing power, i.e., the inability to combine all the elements in a program and make sense of its execution in one's head. Even small pieces of code can be confusing for developers~\cite{Ajami:2017:SPI,DBLP:conf/sigsoft/GopsteinIYDZYC17} at first glance.

For instance, consider the pair of JavaScript code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code snippet on the left-hand side is arguably harder to understand than the one on the right-hand side. The former uses the logical AND operator (\texttt{\&\&}) beyond its lexical meaning~\cite{castor2018} to determine the control flow of the program. Furthermore, the expression has a potential side effect, due to the use of the post-increment operator (\texttt{s++}). The snippet on the right-hand side uses an \texttt{if} statement for control flow and increments the value of variable \texttt{s} using an assignment, making it easier to understand. Notwithstanding, confusion is a direct consequence of the experience of the person who is reading the code. A developer who customarily employs the idiom on the left-hand side may find it easier to understand.

\begin{figure*}[thb]
\noindent\begin{minipage}{.48\textwidth}  
\begin{lstlisting}[language=JavaScript]
res = ps.reduce((s, b) => {
  b && s++
  return s
}
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.48\textwidth}
\begin{lstlisting}[language=JavaScript]
res = ps.reduce((s, b) => {
  if(b) {
    s = s + 1
  }
  return s
}
\end{lstlisting}
\end{minipage}
\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project (left-hand side) and an alternative implementation (right-hand size).}\label{fig:lst01}
% On the right-hand side, we present an alternative implementation without atoms of confusion. We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p.length})).}
\end{figure*}

%Line 2 on the left-hand side of the figure contains two atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic as Control Flow and Post-increment.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ,TheEyesDoNotLie} has attempted to elicit and catalog simple code patterns and language constructs that tend to confuse developers when reading them and for which there are less confusing, functionally-equivalent alternatives. These confusing constructs and patterns are called ``atoms of confusion''~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} when it is possible to experimentally ascertain that there is a less confusing alternative (otherwise they are called ``atom candidates''). The snippet on the left-hand side of Figure~\ref{fig:lst01} contains two atoms of confusion discussed in previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} focusing on the \clang language: Logic as Control Flow and Post-increment. Previous work has identified atoms of confusion in two languages, \clang and Java. According to the studies of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment from small \clang code snippets improved the ability of study participants to predict their outcomes by 41\% and 34\%, respectively. For small Java code snippets, Langhout and Aniche~\cite{Langhout:2021:ACJ} report improvements of 53\% and 46.27\%, respectively. These results show that atoms of confusion may be \textbf{challenging} for developers using these languages. In addition, a study~\cite{DBLP:conf/msr/GopsteinZFC18} with 14 large-scale open source projects written in \clang found that 4.38\% of the lines of code have an atom and their presence has a strong correlation with bug fixing commits and long code comments. These results show that atoms of confusion are \textbf{prevalent} in large systems. They are also \textbf{relevant}, even for experienced software developers. 
%According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.

%factors such as the wrong choice of names for variables and methods~\cite{avidan:icpc2017} and the use of uncommon language constructs~\cite{VEM2018} may confused developers might hinder developers from understanding a piece of code. Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings. That is the case of atoms of confusion: small pieces of code that prevent developers from program comprehension, even though functionally equivalent alternatives exist and are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.

The goal of this paper is twofold. First, it aims to investigate whether atoms of confusion that were identified in Java and C also cause confusion to JavaScript developers. The latter, in spite of the syntactic similarities to Java, is a dynamically-typed language that is arguably more related to Scheme than it is to Java~\cite{Eich:2018:BHJ}. \revised{Previous work has shown that, even for similar languages such as C\# and Java, programmer behavior may differ significantly~\cite{Cabral:2007:EHF,Ebad:2016:ECJ}.}
% This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript. In this sense, because 
JavaScript's programming culture differs from that of languages such as \clang and Java, particularly due to its dynamic capabilities and weaker type system. \revised{Furthermore, it includes elements such as array and object destructuring and spread that have no parallel in Java or C.
At the same time, the language is not so different that many} previously identified atoms cannot be represented in it. Second, the paper aims to verify whether other code patterns, not investigated in previous work, also tend to confuse JavaScript developers. Some of these code patterns are specific to JavaScript, e.g., object destructuring. 
% \castor{The previous sentence is really bad.} 
%Second, we want to analyze whether these atoms also cause confusion when we control for developer experience. As mentioned before, experience seems to play a significant role in whether a certain construct or code pattern can be considered an atom of confusion or not. Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} report that, in their studies, experience is positively correlated with better results. In this paper, we aim to use an experimental design that controls for experience, so as to have greater confidence in our results. 
%, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. 
%Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, 
%JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.\castor{I think we should only say this if we explicitly point out examples of such features}. 
%, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional,
% and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time
% evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.

\revised{We present the results of a mixed-method research effort based on a repository mining study, two experiments, and a set of interviews. 
%, and a mining software repositories study.
The repository mining study measures the prevalence of the atom candidates and helps us motivate the importance of studying these code patterns in more detail. We mined popular JavaScript projects from GitHub looking for the atom candidates that were investigated in both experiments. We found out 
%We found out that atoms frequently appear in practice. Finally, 
%our MSR effort reveals 
that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.} 
The two experiments ask the participants to predict the output of small code snippets, where some of them have atom candidates and others do not. These two experiments were designed and conducted independently, by different researchers, with different samples, and complementary methodologies; one uses a repeated measures design~\cite{Keselman:2001:ARM} whereas the other uses a Latin square design~\cite{Hunter-Experimenters}.  
Both our experiments suggest that two code patterns that have been previously observed to confuse C programmers also confuse JavaScript programmers (with statistical significance, after correction): the comma operator and assignments being used as values. In addition, pre- and post-increment operators, when used as values, omitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one of the two experiments. For all these cases effect sizes were either medium or high. In addition, we have discovered that object destructuring and automatic semicolon insertion are potential JavaScript-specific atoms that have not been discussed in previous work.

To learn more about the misunderstandings in the snippets used in the experiments, we also interviewed 15 experienced professional developers. Participants preferred the code snippets without the atoms in 70\% of the cases. The frequency with which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code.

%% The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). For 5 of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for 4 atoms and medium for 1 of them. 
%% The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
%% To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants preferred the code snippets without the atoms in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.

%% The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code. To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms frequently appear in practice. 
%% Our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects and two of them, Implicit Predicate and Ternary Operator, occur more than once for every 100 lines of code in the dataset. 
%Indeed, four atom candidates are frequently used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

%\castor{I would rather remove the following. We don't even talk about the "library and infrastructure".}
%Altogether, the main contributions of this paper are:

%\begin{itemize}

%\item A mixed-method research effort---based on two experiments, interviews, and a software
%  repository mining study---about the impact of atoms of confusion on JavaScript programs.
  % (a dynamic language), controlling for subject experience.

%  \item A comparison of the impact and prevalence of atoms of confusion in
%    JavaScript with what has been previously reported for other languages
%    (\clang, \cpplang, and Java).
%    \item A library and infrastructure to mine atoms of confusion using 
%    code queries (implemented in the .QL language). This library and 
%    infrastructure are publicly available and might help researchers and 
%    practitioners to automate program analysis at a large scale.

%\end{itemize}

%As implications, our results can help researchers and practitioners to better understand atoms of confusion in JavaScript. For example, our findings might alert the JavaScript community to avoid writing code with certain atoms of confusion. Moreover, our mixed study might help researchers and practitioners to design program transformation tools to remove atoms of confusion in JavaScript.

%They might prioritize atoms that cause a lot of confusion and are very common in the repositories, for instance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FINAL INTRODUCAO - MARCIO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript developers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The source code of a program can be naively understood as a collection of sentences in a language that would be interpreted by a computer (or compiled into a lower level representation and then executed by a computer). Nonetheless, this understanding is far from being complete. While it is not possible to execute programs without writing them in a way that the machine is able to interpret, it is also very difficult to develop software that performs complex tasks if developers do not care about how easily other programmers are going to understand its source code \cite{DBLP:journals/cj/Knuth84}.

% \rb{acho que esse proximo paragrafo pode ser excluido}
% \diego{também acho}

% {\color{blue}That means programming should, in various ways, be regarded as an act of communication. In this context, the programmer is often faced with a dilemma. When initially confronted with a problem they have to solve, the programmer will proceed to employ their cognition into developing the logic of the solution, whilst simultaneously having to worry about writing ``syntactically correct'' code for the machine to parse and translate. The problem is that, during the time spent solving the problem and communicating with the computer, developers often do not have enough cognitive resources to also make considerations about whether their code is understandable by other programmers. Therefore, it takes a change of track to make the code easier to understand \cite{DBLP:books/daglib/0019908}. Not proceeding with such change of track is one of the reasons why ``confusing'' language constructs and idioms arise and remain in a program.}

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Source code misunderstandings can occur due to diverse reasons. For instance, a poorly understood requirement might lead to poorly implemented code. An intricate algorithm may require extra time to be fully comprehended. Besides that, programming languages can also be at fault and specific syntax and semantics of a language can cause misunderstandings. The same applies to some code idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.

% \diego{acho que o prox. parágrafo deve ser adaptado ou até removido}
% {\color{blue}A direct consequence of the presence of these idioms in the code base is that it becomes even more difficult to develop new features for the software. 
% According to Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} ``\emph{the ability to understand pre-existing source code is one of the most important elements of a continuously successful software project}'', it follows that, if one cannot properly understand the \emph{purpose} and \emph{meaning} of a piece of code, they are not going to be able to build upon its existing parts.
% Other consequences of the presence of confusing code include: 
% (a) increased probability of bug introduction, due to a lack of proper understanding of what the program does; 
% and (b) loss of internal quality measures, such as cohesion and coupling. 
% This phenomenon usually leads to an architectural erosion. 
% All these situations, which usually compound each other, expose the necessity of developing recommendations, tools, and techniques that help developers to avoid writing confusing code, identify potentially confusing code, and transform confusing code into clear code. 
% Therefore, it is necessary to understand what programming language constructs and idioms might lead to code that is hard to understand.}

%This research focuses on existing JavaScript idioms that a previous research work have proved to introduce misunderstanding in \clang.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations. All these features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer. To achieve that, we investigate JavaScript idioms that might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in JavaScript---that is, small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. On the right-hand side, we present a clean version of the code, that is, without atoms of confusion. In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability. However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system. Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript developers.

% \begin{figure*}[thb]
% \noindent\begin{minipage}{.45\textwidth}  
% \begin{lstlisting}[language=JavaScript]
% res = properties.reduce((s, b) => {
%   b && s++
%   return s
% }
% \end{lstlisting}
% \end{minipage}\hfill
% \begin{minipage}{.45\textwidth}
% \begin{lstlisting}[language=JavaScript]
% res = properties.reduce((s, b) => {
%   if(b) {
%     s = s + 1
%   }
%   return s
% }
% \end{lstlisting}
% \end{minipage}
% \captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
%   atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic
%   as Control Flow} and \emph{Post-increment}. On the right-hand side, we present an alternative implementation without atoms of confusion. We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p).length}).}\label{fig:lst01}
% \end{figure*}

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 JavaScript developers. In the survey,  their task was to predict the output of short code snippets, with half of them containing atoms of confusion. {\color{red}The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases  the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers. Participants considered the code snippets without the atom candidates easier to understand in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms that might lead developers to misunderstand JavaScript code. The occurrence of atoms in open-source repositories might reveal cases where developers should be more or less careful when writing code---and open the possibility to implement new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects from GitHub and found four atom candidates that are highly used in JavaScript code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Altogether, the main contributions of this paper are: \rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}

% \begin{itemize}
%     \item A study about the impact of atoms of confusion on a
%     dynamic language.
%     \item A list of additional atoms of confusion addressing particular 
%     JavaScript constructs and idioms.
%     \item A library and infrastructure to mine atoms of confusion using 
%     code queries (implemented in the .QL language). This library and 
%     infrastructure are freely available and might help researchers and 
%     practitioners to automate program analysis at a large scale.
% \end{itemize}

% Our research aims to identify and characterize JavaScript \textit{atoms of confusion}: small pieces of code that hinder developers' understanding of the source code of a program~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}.
% Previous research has confirmed the negative impact of atoms of confusion when considering statically typed languages (e.g., \clang and \cpplang)~\cite{DBLP:journals/ese/MedeirosLAAKRG19, DBLP:conf/msr/GopsteinZFC18}.

% Recent works have explored the impact of \emph{atoms of confusion}~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/He19,DBLP:conf/msr/GopsteinZFC18} on program understanding.
% Atoms of confusion correspond to small pieces of code that hinder developers' understanding of the source code of a program. 
% Previous research has confirmed the negative impact of atoms of confusion when considering statically typed languages (e.g., \clang and \cpplang)~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/msr/GopsteinZFC18}.

% In this work, to the best of our knowledge, we are the first to explore this issue in the context of a dynamically typed language (JavaScript), answering the following research questions:
% \emph{How do JavaScript developers  identify atoms of confusion as contributing to program misunderstanding?}
% \emph{What is the impact of atoms of confusion on misunderstanding JavaScript code?},
% and \emph{What is the frequency of occurrence of  atoms of confusion in practice?} 
% To answer these research questions, we conduct this investigation using a mixed-methods approach, involving a survey, a set of interviews, and an activity of mining source code repositories. Altogether, the main contributions of this paper are

% \rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}

% \begin{itemize}
%     \item A first study about the impact of atoms of confusion on a
%     dynamic language. {\color{red}present some of the results here.}
%     \item A library and infrastructure to mine atoms of confusion using 
%     code queries (implemented in the .QL language). This library and 
%     infrastructure is freely available and might help other researcher and 
%     practitioners to automate program analysis at a large scale. 
%     \item A list of additional atoms of confusion addressing particular 
%     JavaScript constructs and idioms. {\color{red}present some insights 
%     about these additional atoms here}
% \end{itemize}